### 2017-02-04 abstract.tex
fresh changes

\begin{abstract}
The term Approximate Dynamic Programming (ADP) refers to a gamut of approximate solution methods for MDPs with large number of states. Though various ADP algorithms are known till date, very few of them successfully address both the prediction and the control problems. Approximate Linear Programming (ALP) is an ADP method that offers sound theoretical guarantees and solves both the prediction and the control problems. Nevertheless, ALP has a serious limitation in that it has large number of constraints, and in practice, a reduced linear program (RLP) is solved instead. Though the RLP has been shown to perform well empirically, error bounds are available only for a specific RLP obtained under idealized assumptions.\\
In this paper, we generalize the RLP to define a generalized reduced linear program (GRLP) which has a tractable number of constraints that are obtained as positive linear combinations of the original constraints of the ALP. The main contribution of this paper is the novel theoretical framework developed to obtain error bounds for any given GRLP. Central to our framework are two $\max$-norm contraction operators. Our result also theoretically justifies linear approximation of constraints. We discuss the implication of our results in the contexts of ADP and reinforcement learning.
\end{abstract}
\begin{keywords}{
Approximate Dynamic Programming (ADP), Markov Decision Processes (MDPs), Approximate Linear Programming (ALP), Generalized Reduced Linear Program (GRLP), Constraint Sampling, Reinforcement Learning.}
\end{keywords}

>>>>

\begin{abstract}
Approximate linear programming (ALP) and its variants have been widely applied to Markov Decision Processes (MDPs) with a large number of states. A serious limitation of ALP is that it has an intractable number of constraints, as a result of which constraint approximations are of interest. In this paper, we define a generalized reduced linear program (GRLP) that has a tractable number of constraints, which are obtained as positive linear combinations of the original constraints of the ALP. The main contribution of this paper is a novel theoretical framework developed to obtain error bounds for any given GRLP. By providing a detailed error analysis for the GRLP, we justify linear approximation of the dual variables. Unlike prior results on constraint sampling, our analysis is deterministic and is based on a novel contraction operator.
\end{abstract}
\begin{keywords}{
Approximate Dynamic Programming (ADP), Markov Decision Processes (MDPs), Approximate Linear Programming (ALP), Generalized Reduced Linear Program (GRLP), Constraint Sampling, Reinforcement Learning.}
\end{keywords}

### 2017-02-04 adp.tex
fresh changes

%!TEX root =  autocontgrlp.tex
\section{Approximate Dynamic Programming}
In practical problems, often the states are best described as multi-dimensional tuples.
In this context, the term \emph{curse-of-dimensionality} refers to the fact that the number of states
	grows exponentially as a function of the number of dimensions
	and thus solution methods whose computation cost grows at least linearly with the size of the state space
	quickly become intractable.
%Most MDPs arising in practical applications suffer from the curse, i.e., have large number of states and it is difficult to compute $J^*/J_u\in \R^n$ exactly in such scenarios. 
Approximate dynamic programming (ADP) \cite{lspi,lspe,ALP,wang2014approximate} methods compute an approximate value function $\tilde{J}$ instead of $J^*/J_u$. 
In order to tackle the curse, ADP methods resort to dimensionality reduction by parametrizing the value function and/or the policy. Value-function based ADP methods form an important subclass of ADP methods whose brief overview is given below.

@@

In the next section, we discuss the approximate linear programming (ALP) formulation, the basic results and present prior results in literature as well as motivate the problem that we address in this paper.

>>>>

%!TEX root =  autocontgrlp.tex
\section{Approximate Dynamic Programming: Objectives and Methods}
\FloatBarrier
\begin{table}[H]
\resizebox{\columnwidth}{!}{
\begin{tabular}{|c|c|c|}\hline
ADP Method	&Empirical	&Theoretical\\\hline
Projected Bellman &\ding{51}    &\ding{53}-Policy Chattering\\
Equation	&	&\\\hline
ALP		&\ding{53}-Large number of Constraints &\ding{51}\\\hline
RLP		&\ding{51} &\ding{53}-  Only under\\
&&ideal assumptions\\\hline
\end{tabular}
}
\end{table}

\begin{comment}
In practical problems, often the states are best described as multi-dimensional tuples.
In this context, the term \emph{curse-of-dimensionality} refers to the fact that the number of states
	grows exponentially as a function of the number of dimensions
	and thus solution methods whose computation cost grows at least linearly with the size of the state space
	quickly become intractable.
%Most MDPs arising in practical applications suffer from the curse, i.e., have large number of states and it is difficult to compute $J^*/J_u\in \R^n$ exactly in such scenarios. 
Approximate dynamic programming (ADP) \cite{lspi,lspe,ALP,wang2014approximate} methods compute an approximate value function $\tilde{J}$ instead of $J^*/J_u$. 
In order to tackle the curse, ADP methods resort to dimensionality reduction by parametrizing the value function and/or the policy. Value-function based ADP methods form an important subclass of ADP methods whose brief overview is given below.

@@

In the next section, we discuss the approximate linear programming (ALP) formulation, the basic results and present prior results in literature as well as motivate the problem that we address in this paper. 
\end{comment}

### 2017-02-04 errana.tex
fresh changes

%!TEX root =  autocontgrlp.tex
\section{Error Analysis}\label{sec:errana}
\subsection{Premilinary results from MDP}
We now state without proof the most important properties of the Bellman operator(s).
The proofs are immediate from the definitions, but can also be found in \cite{BertB}.
\begin{comment}
First, we introduce some extra notation:
For $J_1,J_2\in \R^n$, we write $J_1\le J_2$ if $J_1(s)\le J_2(s)$ holds for all $s\in S$.
We use $\one \in \R^n$ to denote a vector with all entries $1$.
The maximum norm $\norm{\cdot}_{\infty}$ is defined by $ \norm{v}_{\infty} = \max_{s\in S} |v(s)|$.
\end{comment}
\begin{lemma}\label{tprop}
\begin{enumerate}[(i)]
\item \label{monotone}$T$ is a monotone map, i.e., given $J_1,J_2 \in \R^n$ such that $J_1\leq J_2$, we have $T J_1\leq T J_2$.
\item \label{shift}
Given $J\in \R^n$ and $t \in \R$, we have $T(J+t\one)=TJ+\alpha t\one$.
\item \label{maxnorm}
If $T: \R^n \to \R^n$ is any operator that is monotonous and satisfies~\eqref{shift} then
$T$ is a $\max$-norm contraction operator with contraction factor $\alpha \in (0,1)$, i.e., given $J_1, J_2 \in \R^n$,
$
||TJ_1-TJ_2||_\infty\leq \alpha ||J_1-J_2||_\infty.
$
\item \label{uniquesol}
$J^*$ is a unique fixed point of $T$, i.e., $J^*=TJ^*$.
\end{enumerate}
\end{lemma}
\begin{corollary}
If $J\in \R^n$ is such that $J\geq TJ$ then $J\geq TJ^2\geq \ldots \geq J^*$.
\end{corollary}
Though \cref{tprop} are stated for the Bellman operator $T$, the results also hold for $H$ as well.
\subsection{Analysis using Lyapunov Functions}
\begin{definition}
Let $c,\rho,\chi:S \to \R_+$ be positive valued functions, where $\R_+$ denotes the set of strictly positive reals. Then for $J\in \R^n$, $a\in A$ and $s\in S$, define
\begin{enumerate}[(i)]
\item The discounted maximal inflation of $\chi$ due to $P = (p_a)_{a\in A}$ as $\beta_{\chi}=\max_{s \in S} \frac{\underset{a \in A}{\max}\big(\alpha\sum_{s'}p_a(s,s')\chi(s')\big)}{\chi(s)}$.
\item The function $\chi:S\to\R_+$ above to be a \emph{Lyapunov} function for $P = (p_a)_{a\in A}$ if $\beta_{\chi}<1$.
\end{enumerate}
\end{definition}
\begin{assumption}\label{grlpassmp}
\begin{enumerate}[(i)]
%\item \label{one} The first column of the feature matrix $\Phi$ (i.e., $\phi_1$) is $\one \in \R^n$.
\item \label{lyap} $\psi\colon S \ra \R_+$ is a Lyapunov function for $P$
and is present in the column span of the feature matrix $\Phi$: For some $r_0\in \R^k$, $\Phi r_0 = \psi$.
\item \label{ass:n4} The set $\N'$ is such that $\N' = \N + t r_0$ for any $t\in \R$, where $r_0\in \R^k$ such that $\Phi r_0 = \psi$.
\end{enumerate}
\end{assumption}
The error bounds for the ALP in \cite{ALP} are expressed in terms of $\frac{1}{1-\beta_{\psi}}$.  A smaller $\beta_{\psi}$ loosely implies `stability' of the the underlying MDP,  with smaller values representative of higher stability. Prior works in ALP literature \cite{ALP,SALP,CS} make use of Lyapunov function based analysis to obtain error bounds that exploit the structure of the underlying MDP. In particular, the prior bounds suggests that the ALP is likely to generate good approximations when the underlying MDP is stable. We also adopt a similar approach by stating our results using Lyapunov function based arguments.

@@

\label{eq:psilin}
T(J+ t \psi ) \le TJ + \beta_{\psi}\,t\,  \psi\,. %\,\, \quad (J\in \R^n,\, t>0)\,.

@@

Given $J\in \R^n$, let $\F_J\eqdef\{\,\Phi r\,: W^\top E \Phi r\geq W^\top HJ, r\in \N'\,\}\,$. Choose any $i\in \{1, \ldots, n\}$. Since $J_1\leq J_2$, from \cref{monotone} and \cref{grlpassmp}-\eqref{wassump} it follows that $W^\top H J_1\leq W^\top H J_2$. Hence, $\F_{J_2} \subset \F_{J_1}$ and thus $(\hg J_1)(i) \le (\hg J_2)(i)$.  Since $i$ was arbitrary, the result follows. 

@@

\hg( J + t \psi) \le \hg J + \beta t \psi,

@@

for any $J\in \R^n$ and $t\ge 0$.

@@

Now, using Lemma~\ref{lpsol} with $A=W^\top E \Phi$, $b=W^\top HJ$, $d=e_i\Phi$, $b_0=t\beta_\psi \psi$

@@

%The result follows in a similar manner as the proofs for Lemmas~\ref{gshift} and ~\ref{tgshift} by using the result in Lemma~\ref{lpsol}.

@@

Hence, by the monotone property of $H$ and \cref{grlpassmp}-\eqref{wassump} it is sufficient if $\Phi r_i \ge V_0$.

@@

We know from Lemma~\ref{relation2} that $\hj = \Phi \hr \geq\hv$, and thus

@@

We know from Lemma~\ref{relation2} that $\tj = \Phi \tr \geq\hv$.

@@

Then, by choosing $r'$ as in Lemma~\ref{restate} we have

@@

Now, $r'$ is feasible for the ALP in \eqref{grlp} by \cref{restate}. 

@@

The result now follows from Lemma~\ref{cmt1mn} and Theorem~\ref{mt2mn}.

@@

>>>>

%!TEX root =  autocontgrlp.tex
\section{Error Analysis}\label{sec:improv}
\subsection{Properties of the Bellman Operator}
We now state without proof the most important properties of the Bellman operator(s).
The proofs are immediate from the definitions, but can also be found in \cite{BertB}.
\begin{comment}
First, we introduce some extra notation:
For $J_1,J_2\in \R^n$, we write $J_1\le J_2$ if $J_1(s)\le J_2(s)$ holds for all $s\in S$.
We use $\one \in \R^n$ to denote a vector with all entries $1$.
The maximum norm $\norm{\cdot}_{\infty}$ is defined by $ \norm{v}_{\infty} = \max_{s\in S} |v(s)|$.
\end{comment}
\begin{lemma}\label{tprop}
\begin{enumerate}[(i)]
\item \label{monotone}$T$ is a monotone map, i.e., given $J_1,J_2 \in \R^n$ such that $J_1\leq J_2$, we have $T J_1\leq T J_2$.
\item \label{shift}
Given $J\in \R^n$ and $t \in \R$, we have $T(J+t\one)=TJ+\alpha t\one$.
\item \label{maxnorm}
If $T: \R^n \to \R^n$ is any operator that is monotonous and satisfies~\eqref{shift} then
$T$ is a $\max$-norm contraction operator with contraction factor $\alpha \in (0,1)$, i.e., given $J_1, J_2 \in \R^n$,
$
||TJ_1-TJ_2||_\infty\leq \alpha ||J_1-J_2||_\infty.
$
\item \label{uniquesol}
$J^*$ is a unique fixed point of $T$, i.e., $J^*=TJ^*$.
\end{enumerate}
\end{lemma}
\begin{corollary}
If $J\in \R^n$ is such that $J\geq TJ$ then $J\geq TJ^2\geq \ldots \geq J^*$.
\end{corollary}
Though \cref{tprop} are stated for the Bellman operator $T$, the results also hold for $H$ as well.\par
We now present the analysis to derive the improved bounds where the idea is to show that the novel projection operators ($\Gamma/\hg$) are contraction maps. To this end, we go through steps similar to \Cref{tprop}-\eqref{monotone},~\eqref{shift} and ~\eqref{maxnorm}. Much the results that ensue are based on `Lyapunov' analysis where the idea is to replace the constant function $\one$ by a certain Lyapunov function (see \Cref{def:lyap}) and the $\norm{\cdot}_{\infty}$ by a weighted max-norm (see \Cref{notations}-\eqref{norms}).
\subsection{Analysis using Lyapunov Functions}
\begin{definition}\label{def:lyap}
Let $c,\rho,\chi:S \to \R_+$ be positive valued functions, where $\R_+$ denotes the set of strictly positive reals. Then for $J\in \R^n$, $a\in A$ and $s\in S$, define
\begin{enumerate}[(i)]
\item The discounted maximal inflation of $\chi$ due to $P = (p_a)_{a\in A}$ as $\beta_{\chi}=\max_{s \in S} \frac{\underset{a \in A}{\max}\big(\alpha\sum_{s'}p_a(s,s')\chi(s')\big)}{\chi(s)}$.
\item The function $\chi:S\to\R_+$ above to be a \emph{Lyapunov} function for $P = (p_a)_{a\in A}$ if $\beta_{\chi}<1$.
\end{enumerate}
\end{definition}
\begin{assumption}\label{grlpassmp}
\begin{enumerate}[(i)]
%\item \label{one} The first column of the feature matrix $\Phi$ (i.e., $\phi_1$) is $\one \in \R^n$.
\item \label{lyap} $\psi\colon S \ra \R_+$ is a Lyapunov function for $P$
and is present in the column span of the feature matrix $\Phi$: For some $r_0\in \R^k$, $\Phi r_0 = \psi$.
\item \label{ass:n4} The set $\N'$ is such that $\N' = \N + t r_0$ for any $t\in \R$, where $r_0\in \R^k$ such that $\Phi r_0 = \psi$.
\item \label{wassmp} $W \in \R^{nd\times m}_+$ is a matrix with all positive entries.
\end{enumerate}
\end{assumption}
The error bounds for the ALP in \cite{ALP} are expressed in terms of $\frac{1}{1-\beta_{\psi}}$.  A smaller $\beta_{\psi}$ loosely implies `stability' of the the underlying MDP,  with smaller values representative of higher stability. Prior works in ALP literature \cite{ALP,SALP,CS} make use of Lyapunov function based analysis to obtain error bounds that exploit the structure of the underlying MDP. In particular, the prior bounds suggests that the ALP is likely to generate good approximations when the underlying MDP is stable. We also adopt a similar approach by stating our results using Lyapunov function based arguments.

@@

\begin{align}\label{eq:psilin}
T(J+ t \psi ) \le TJ + \beta_{\psi}\,t\,  \psi,\\
H(J+ t \psi ) \le HJ + E \beta_{\psi}\,t\,  \psi\,.

@@

Given $J\in \R^n$, let $\F_J\eqdef\{\,\Phi r\,: W^\top E \Phi r\geq W^\top HJ, r\in \N'\,\}\,$. Choose any $i\in \{1, \ldots, n\}$. Since $J_1\leq J_2$, from \Cref{tprop}-\eqref{monotone} and \cref{grlpassmp}-\eqref{wassmp} it follows that $W^\top H J_1\leq W^\top H J_2$. Hence, $\F_{J_2} \subset \F_{J_1}$ and thus $(\hg J_1)(i) \le (\hg J_2)(i)$.  Since $i$ was arbitrary, the result follows.

@@

\hg( J + t \psi) \le \hg J + \beta t \psi.

@@

@@

Now, using \Cref{lpsol} with $A=W^\top E \Phi$, $b=W^\top HJ$, $d=e_i\Phi$, $b_0=t\beta_\psi \psi$

@@

@@

Hence, by the monotone property of $H$ and \cref{grlpassmp}-\eqref{wassmp} it is sufficient if $\Phi r_i \ge V_0$.

@@

We know from \Cref{relation1} that $\hj = \Phi \hr \geq\hv$, and thus

@@

We know from \Cref{relation2} that $\tj = \Phi \tr \geq\hv$.

@@

Then, by choosing $r'$ as in \Cref{restate} we have

@@

Now, $r'$ is feasible for the ALP in \eqref{alp} by \cref{restate}.

@@

The result now follows from \Cref{cmt1mn} and \Cref{mt2mn}.

@@

\begin{theorem}[Constraint Sampling]
Let $s\in S$ be a state whose constraint is selected by $W$ (i.e., for some $i$ and all $(s',a)\in S\times A$,
$W_{s'a,i}=\delta_{s=s'}$.
Then
$
|\Gamma J^*(s)-\hg J^*(s)|<|\Gamma J^*(s)-J^*(s)|.
$

\begin{proof}
Let $r_{e_s,J^*}$ and ${r}'_{e_s,J^*}$ be solutions to the linear programs in \eqref{lubplp} and \eqref{alubplp} respectively for $c=e_s$ and $J=J^*$. It is easy to note that $r_{e_s,J^*}$ is feasible for the linear program in \eqref{alubplp} for $c=e_s$ and $J^*$, and hence it follows that $(\Phi r_{e_s,J^*})(s)\geq (\Phi {r}'_{e_s,J^*})(s)$. However, since the constraints with respect to state $s$ have been chosen we know that $(\Phi {r}'_{e_s,J^*})(s)\geq J^*(s)$. The proof follows from noting that $(\Gamma J^*)(s)=(\Phi r_{e_s,J^*})(s)$ and $\hg J^*(s)=(\Phi {r}_{e_s,J^*})(s)$.
\end{proof}

### 2017-02-04 grlp.tex
fresh changes

%!TEX root =  autocontgrlp.tex
\section{Approximate Linear Programming: Successes and Challanges}
When the MDP has a large number of states, it is difficult to solve for $J^*$ using either the linear program \eqref{mdplp} or other full state representation methods such as value iteration or policy iteration \cite{BertB}. A practical solution is to resort to function approximation. Linear function approximation, wherein the solution is searched in the subspace spanned by the column vectors of a given feature matrix $\Phi$.
The approximate linear program (ALP) is obtained by making use of LFA in the LP, i.e., by introducing the new variables $r\in \R^k$ and adding the extra constraint $J=\Phi r$ in \eqref{mdplp} with $\Phi \in \R^{n\times k}$ \citep{SchSei85}.
By substitution, this leads to
\begin{align}\label{alp}
\min_{r\in \R^k}\, &c^\top \Phi r\,\,\,
\text{s.t.}\mb \Phi r\geq T \Phi r,
where $J\geq TJ$ is a shorthand for the $nd$ constraints in \eqref{mdplp} and $\Phi$ is a feature matrix whose first column is $\one$. Unless specified otherwise we use $\tr$ to denote an arbitrary solution to the ALP, and we let $\tj=\Phi \tr$ to denote the corresponding approximate value function and $\tu$ to denote the greedy policy w.r.t. $\tj$.
The following is a preliminary error bound for the ALP from \cite{ALP}:
\begin{theorem}[Error Bound for ALP]
||J^*-\tj||_{1,c}\leq \frac{2}{1-\alpha}\min_{r}||J^*-\Phi r||_\infty

@@

If $\tilde{r}_{RLP}$ is a solution to the RLP, then

@@

@@

@@

where we assume that $W \in \R^{nd\times m}_+$ is a matrix with all positive entries, $\N$ is an additional (compact) constraint set to ensure the boundedness of the solution and $\tr \in \N$.
\footnote{The appendix explains how $\N$ can be chosen.}
In what follows, we denote the solution to the GRLP by $\hr$, the approximate value function by $\hj=\Phi \hr$ and use $\hu$ to denote the greedy policy w.r.t. $\hj$.\par
%\subsection{Novel Projection Operators}
We now define two projection operators which are central to our error analysis and in them we assume that the set $\N'\subset \R^k$ is such that $\N' = \N + t \one$ for any $t\in \R$.

@@

%The approximate least upper bound (ALUB) projection operator $\hg \colon \R^n \ra \R^n$ is defined as below:
The definition of the second operator is as follows:
\begin{definition}\label{alubpop}
Given $J\in \R^n$ and the nonnegative valued vector $c\in \R^n_+$, define $r'_{c,J}$ to be the solution to
\begin{align}\label{alubplp}
\underset{r\in \N'}{\min}& \,\mb c^\top \Phi r\,\mb
\text{s.t.} \,\,\, W^\top E \Phi r\geq W^\top HJ.
\end{align}
Then, the \emph{approximate least upper bound (ALUB) projection operator}
$\hg \colon \R^n \ra \R^n$ is defined as
\begin{align}\label{tgamdef}
(\hg J)(i)\eqdef(\Phi r'_{e_i,J})(i), \mb i=1,\ldots,n\,, J\in \R^n\,.
\end{align}
\end{definition}
\begin{remark}\label{ubrem}
To understand the meaning of $\Gamma$ (and $\hg$) define
\begin{align}\label{ubclass}
\F_J\eqdef\{\,\Phi r\,:\,\Phi r\geq TJ, r\in \N\,\},
\end{align}
where $J\in \R^n$.
Disregarding the constraint $r\in \N$,
$\F_J$ contains vectors in the span of $\Phi$ that upper bound $TJ$. Further, since $(\Gamma J)(i) = \min\{ V(i) \,:\, V\in \F_J \}$, it also follows that $ V\ge \Gamma J $ holds for any $V\in \F_J$.
\end{remark}

%\input{cartoon}
\subsection{Error Bounds}
\begin{theorem}[Error Bound for GRLP]
\label{cmt2}
%Let $\hj$, $\hv$, $r^*$ and $J^*$ be as in Theorem~\ref{mt2mn}, then
It holds that
\begin{align}\label{finalbnd}
\begin{split}
||J^*-\hj||_{1,c}
\leq\frac{1}{1-\alpha}(6 ||J^*-\Phi r^*||_{\infty}
+2||\Gamma J^*-\hg J^*||_{\infty}).
\end{split}
\end{align}
\end{theorem}
\begin{comment}
\begin{theorem}[Control Error Bound in $\norm{\cdot}_{\infty}$]
\label{polthe}
Let $\hu$ be the greedy policy with respect to the solution $\hj$ of the GRLP and $J_{\hu}$ be its value function.
% Let $r^*$ be as in Theorem~\ref{mt2mn}, then
Then,
\begin{align}\label{polthebnd}
%||J_{\hu}-\hj||_{1,c}
\norm{J^* - J_{\hu}}_{1,c}
&\leq 2\left(\frac{c^\top \psi}{(1-\beta_{\psi})^2}\right)\, \big( 2||J^*-\Phi r^*||_{\infty}
\nn\\&
+||\Gamma J^*-\hg J^*||_{\infty}+||\hj-\hg\hj||_{\infty}\big).
\end{align}
\end{theorem}
\end{comment}
\begin{theorem}[Constraint Sampling]\label{st}
Let $s\in S$ be a state whose constraint is selected by $W$ (i.e., for some $i$ and all $(s',a)\in S\times A$,
$W_{s'a,i}=\delta_{s=s'}$.
Then
\begin{align*}%\label{sampexp}
|\Gamma J^*(s)-\hg J^*(s)|<|\Gamma J^*(s)-J^*(s)|.
\end{align*}
\end{theorem}
\begin{comment}
An interesting feature is that unlike prior work on constraint sampling based on concentration inequalities (e.g.,  \cite{CS}), our analysis is based on contraction operators and is completely deterministic.
In particular, the error term $\etmn$ gives new insights into constraint selection:
The result of Theorem~\ref{st} is that
\end{comment}
%The modified $L_\infty$ norm in \eqref{finalbndmn} comes to our rescue to control the error due to those states that are not chosen.

>>>>

%!TEX root =  autocontgrlp.tex
\begin{comment}
\section{Approximate Linear Programming: Successes and Challenges}
When the MDP has a large number of states, it is difficult to solve for $J^*$ using either the linear program \eqref{mdplp} or other full state representation methods such as value iteration or policy iteration \cite{BertB}. A practical solution is to resort to function approximation. Linear function approximation, wherein the solution is searched in the subspace spanned by the column vectors of a given feature matrix $\Phi$.
The approximate linear program (ALP) is obtained by making use of LFA in the LP, i.e., by introducing the new variables $r\in \R^k$ and adding the extra constraint $J=\Phi r$ in \eqref{mdplp} with $\Phi \in \R^{n\times k}$ \citep{SchSei85}.
By substitution, this leads to
\begin{align}\label{alp}
\begin{split}
\min_{r\in \R^k}\, &c^\top \Phi r\,\,\,
\text{s.t.}\mb \Phi r\geq T \Phi r,
\end{split}
\end{align}
where $J\geq TJ$ is a shorthand for the $nd$ constraints in \eqref{mdplp} and $\Phi$ is a feature matrix whose first column is $\one$. Unless specified otherwise we use $\tr$ to denote an arbitrary solution to the ALP, and we let $\tj=\Phi \tr$ to denote the corresponding approximate value function and $\tu$ to denote the greedy policy w.r.t. $\tj$.
The following is a preliminary error bound for the ALP from \cite{ALP}:
\begin{theorem}[Error Bound for ALP]
\begin{align*}
||J^*-\tj||_{1,c}\leq \frac{2}{1-\alpha}\min_{r}||J^*-\Phi r||_\infty
\end{align*}

@@

Let $\mu_{u^*}\eqdef(1-\alpha)c^\top (I-\alpha P_{u^*})^{-1}$ be a probability distribution (of discounted number of visits) over the states $S$. Define $\psi_{u^*}$ to be the distribution amongst state-action pairs such that $\psi(s,a)\eqdef \frac{\mu_{u^*}}{d}$ and define $\theta\eqdef=\frac{1+\alpha}{2 c^\top J^*}\underset_{r\in \N}{\sup}\parallelJ^*-\Phi r \parallel_\infty$. Then for a given $\epsilon>0$ and $\delta>0$ it holds that

@@

for all $m\geq \frac{16d\theta}{(1-\alpha)\epsilon}\big(k\ln\frac{48d\theta}{(1-\alpha)\epsilon}+\ln \frac{2}{\delta}\big)$.

@@

\section{Generalized Reduced Linear Programming}\label{sec:grlp}

@@

where we assume that $W \in \R^{nd\times m}_+$ is a matrix with all positive entries which specifies the linear combination of constraints. The RLP \eqref{rlp} can be recovered as a special case of the GRLP \eqref{grlp} by setting $W$ to the matrix with the $i^{th}$ column having $1$ corresponding to the constraint that was sampled (implicitly assuming that there is a natural ordering of the $nd$ constraints) and all the other entries as $0$. The $\N$ in the RLP \eqref{rlp} is retained in the GRLP \eqref{grlp} as well. For $m=nd$ and $W=I_{nd\times nd}$, the GRLP is same as the ALP. We denote an arbitrary solution to the GRLP by $\hr$, and the approximate value function by $\hj=\Phi \hr$ and use $\hu$ to denote the greedy policy w.r.t. $\hj$.\par
The GRLP introduces linear function approximation in both the primal and dual variables of the LP formulation.
To understand this, we need to look at the Lagrangian of the ALP and GRLP in
\eqref{lag} and \eqref{lag2} respectively, i.e.,
\begin{align}\label{lag}
\tilde{L}(r,\lambda)=c^\top \Phi r+\lambda^\top (T\Phi r-\Phi r), \\ \label{lag2}\hat{L}(r,q)=c^\top \Phi r+q^\top W^\top (T\Phi r-\Phi r).
Thus, when $Wq = \lambda$, i.e., when $W$ is a set of basis functions that allow
a low dimensional linear representation of the dual variables $\lambda$,
the two problems are the same.
%Note that $ Wq\approx \lambda$ in \eqref{lag2} is linear function approximation of the Lagrange multipliers.
Hence, while the ALP employs LFA in its objective function (i.e., use of $\Phi r$), the GRLP employs linear approximation both in the objective function ($\Phi r$) as well as the constraints (use of $W$).
%Further, $W$ can be interpreted as the feature matrix that approximates the Lagrange multipliers as $\lambda\approx Wq$, where $\lambda \in \R^{nd}, r\in \R^m$.
To get a sense of how $W$ should be chosen, recall that
the optimal Lagrange multipliers are the discounted number of visits to the ``state-action pairs'' under an optimal policy $u^*$, i.e.,
\lambda^*(s,u^*(s))&=\big(c^\top(I-\alpha P_{u^*})^{-1}\big)(s)\nn\\
&= \big(c^\top(I+\alpha P_{u^*}+\alpha^2 P_{u^*}^2+\ldots)\big)(s),\nn\\
\lambda^*(s,a)&=0, \qquad \text{for all } a \neq u^*(s),\nn
where $P_{u^*}$ is the probability transition matrix under $u^*$ ($P_{u^*}(s,s') = P_{u^*(s)}(s,s')$, $s,s'\in S$) \cite{dolgov}. Even though we might not have the optimal policy $u^*$ in practice, the fact that $\lambda^*$ is a probability distribution and that it is a linear combination of $\{P_{u^*},P^2_{u^*},\ldots\}$ hints at the kind of features that might be useful for the $W$ matrix.
\FloatBarrier
\input{cartoon}
\Cref{cartoon} which shows the solutions to the LP, ALP and GRLP respectively. The error in the ALP solution has already been studied in \cite{ALP}. Our objective is to study the extra source of error due to constraint approximation. We now define two projection operators which are central to our error analysis and in them we assume that the set $\N'\subset \R^k$ is such that $\N' = \N + t \one$ for any $t\in \R$.
In this section we present the main results of this paper in \Cref{cmt2} (we state improved bounds in \Cref{sec:improv}). Our bounds are expressed in terms of two novel contractions operators which we define in \Cref{lubpop,alubpop}.

@@

The definition of the second operator is as follows:
\begin{definition}\label{alubpop}
Given $J\in \R^n$ and the nonnegative valued vector $c\in \R^n_+$, define $r'_{c,J}$ to be the solution to
\begin{align}\label{alubplp}
\underset{r\in \N'}{\min}& \,\mb c^\top \Phi r\,\mb
\text{s.t.} \,\,\, W^\top E \Phi r\geq W^\top HJ.
Then, the \emph{approximate least upper bound (ALUB) projection operator}
$\hg \colon \R^n \ra \R^n$ is defined as
\begin{align}\label{tgamdef}
(\hg J)(i)\eqdef(\Phi r'_{e_i,J})(i), \mb i=1,\ldots,n\,, J\in \R^n\,.
\begin{remark}\label{ubrem}
To understand the meaning of $\Gamma$ (and $\hg$) define
\begin{align}\label{ubclass}
\F_J\eqdef\{\,\Phi r\,:\,\Phi r\geq TJ, r\in \N\,\},
where $J\in \R^n$.
Disregarding the constraint $r\in \N$,
$\F_J$ contains vectors in the span of $\Phi$ that upper bound $TJ$. Further, since $(\Gamma J)(i) = \min\{ V(i) \,:\, V\in \F_J \}$, it also follows that $ V\ge \Gamma J $ holds for any $V\in \F_J$.
\end{remark}
\subsection{Error Bounds}
\begin{theorem}[Error Bound for GRLP]
\label{cmt2}
Let $\one\in\R^k$ be in the column span of $\Phi$, then it holds that
\begin{enumerate}
\item The prediction error is bound as\\
$||J^*-\hj||_{1,c}
\leq\frac{1}{1-\alpha}(6~\us{\min}{r\in \R^k}||J^*-\Phi r||_{\infty}
+2||\Gamma J^*-\hg J^*||_{\infty})$.
\item The control error is bound as\\
$\norm{J^* - J_{\hu}}_{1,c}
\leq 2\left(\frac{1}{(1-\alpha)^2}\right)\, \big( 2~\us{\min}{r\in \R^k} ||J^*-\Phi r||_{\infty}
+||\Gamma J^*-\hg J^*||_{\infty}+||\hj-\hg\hj||_{\infty}\big)$.
\end{enumerate}
\textbf{On Prediction Error:}The first factor in the right hand side of the prediction error in \Cref{cmt2} is related to the best possible approximation that can be achieved with the chosen feature matrix $\Phi$. This term is inherent to the ALP formulation as shown in \Cref{alpvanilla}. The second factor in the right hand side of the prediction error is related to constraint approximation and is completely defined in terms of $\Phi$, $W$ and $T$, and does not require knowledge of stationary distribution of the optimal policy. It makes intuitive sense since given that $\Phi$ approximates $J^*$, it is enough for $W$ to depend on $\Phi$ and $T$ without any additional requirements.\par
\textbf{On Control Error:} The first two terms are quite similar to those in the bound for prediction error. The third term occurs due to the fact that $\Phi \geq T\Phi r$ that holds in the case of ALP does not hold in the case of GRLP.\par
\begin{theorem}[Control Error Bound in $\norm{\cdot}_{\infty}$]
\label{polthe}
Let $\hu$ be the greedy policy with respect to the solution $\hj$ of the GRLP and $J_{\hu}$ be its value function.
% Let $r^*$ be as in Theorem~\ref{mt2mn}, then
Then,
\begin{align}\label{polthebnd}
%||J_{\hu}-\hj||_{1,c}
\norm{J^* - J_{\hu}}_{1,c}
&\leq 2\left(\frac{c^\top \psi}{(1-\beta_{\psi})^2}\right)\, \big( 2||J^*-\Phi r^*||_{\infty}
\nn\\&
+||\Gamma J^*-\hg J^*||_{\infty}+||\hj-\hg\hj||_{\infty}\big).
\begin{theorem}[Constraint Sampling]\label{st}
Let $s\in S$ be a state whose constraint is selected by $W$ (i.e., for some $i$ and all $(s',a)\in S\times A$,
$W_{s'a,i}=\delta_{s=s'}$.
Then
\begin{align*}%\label{sampexp}
|\Gamma J^*(s)-\hg J^*(s)|<|\Gamma J^*(s)-J^*(s)|.
It is important to note that \Cref{rlpt} holds only in high probability and is valid only under idealized assumption of knowing the optimal policy $u^*$ and \Cref{cmt2} does not have these limitations.
In addition, from \Cref{st}, we know that the additional error $|\Gamma J^*(s) -\hg J^*(s)|$ due to constraint approximation is less than the original projection error $|\Gamma J^*(s)-J^*(s)|$ due to function approximation. This means that for the RLP to perform well it is enough to retain those states for which the linear function approximation via $\Phi$ is known to perform well.
